---
title: "R Notebook"
output: html_notebook
  theme: darkly
---

# Importing the Data

```{r}
library(readxl)
Phen<- read_excel("ThistlePhenology Nov10 for R.xlsx")
Phen$BlockID <- as.factor(Phen$BlockID)
Phen$PlotID <- as.factor(Phen$PlotID)
Phen$Headby7_12 <- as.factor(Phen$Headby7_12)
Phen$FlrBy8_26 <- as.factor(Phen$FlrBy8_26)
Phen$PostAnthBy8_28 <- as.factor(Phen$PostAnthBy8_28)
Phen$X1CensFlrDuration <- as.factor(Phen$`1CensFlrDuration`)
Phen$MaxFlrBy8_26 <- as.factor(Phen$MaxFlrBy8_26)
str(Phen)
```

## Converting Plant Height to Standard Deviation

Will help with convergence, everything is relative

```{r}
mnhgt<-mean(Phen$MaxPlantHgt)
sdhgt<-sd(Phen$MaxPlantHgt)
Phen$standhgt<-(Phen$MaxPlantHgt - mnhgt)/sdhgt
str(Phen)
```

## Investigating how many Individuals in Each Level of Treatment

```{r}
tapply(Phen$Headby7_12,Phen$Fert,length)
tapply(Phen$Headby7_12,Phen$Excl,length)
```

## Investigating how many Y/N in Each Level of Treatment

```{r}
tapply(Phen$Headby7_12,list(Phen$Fert,Phen$Headby7_12),length)
tapply(Phen$Headby7_12,list(Phen$Excl,Phen$Headby7_12),length)
```

# GLMM Preparation

First we want to separate exclusion data into insect exclusion Y/N:

```{r}
library(plyr)
Phen1 <- Phen
Phen1$InsPool <- revalue(Phen$Excl,c("CTL" = "InsPres", "Pseudo" = "InsPres", "TrnchCTL" = "InsPres", "AG" = "InsPres","IN" = "InsAbs", "AGIN" = "InsAbs", "AGINBG" = "InsAbs"))
Phen1$AGPool <- revalue(Phen$Excl,c("CTL" = "AGPres", "Pseudo" = "AGPres", "TrnchCTL" = "AGPres", "IN" = "AGPres",
                                     "AG" = "AGAbs", "AGIN" = "AGAbs", "AGINBG" = "AGAbs"))

Phen1$AGPool <- as.factor(Phen1$AGPool)

Phen1$InsPool <- as.factor(Phen1$InsPool)
```

Then we will examine how Y/N are distributed with ExclPool:

```{r}
tapply(Phen1$PostAnthBy8_28,Phen1$PostAnthBy8_28,length)
```

# GLMMs

```{r}
library(lme4)
```

## Headby7/12 - InsExcl

We would theoretically like the most complex model incorporating both Exclusion and Fertilization but it will most likely not converge, so we will split it.

```{r}
ModelHdByIns<- glmer(Headby7_12~ standhgt +Fert + InsPool +(1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelHdByIns)
drop1(ModelHdByIns, test = "Chi")
```

```{r}

ModelHdByInsN<- glmer(Headby7_12~Fert + InsPool +(1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelHdByInsN)
drop1(ModelHdByInsN, test = "Chi")
```

------------------------------------------------------------------------

## FlrBy8/26 - InsExcl

Was able to obtain the most complex model by removing BlockID as a random effect:

```{r}
ModelFlrByIns<- glmer(FlrBy8_26 ~ InsPool + Fert + standhgt + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelFlrByIns)
drop1(ModelFlrByIns, test = "Chi")
```

```{r}
ModelFlrByInsN<- glmer(FlrBy8_26 ~ InsPool + Fert + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelFlrByInsN)
drop1(ModelFlrByInsN, test = "Chi")
```

------------------------------------------------------------------------

## PostAnthBy 8/28 - InsExcl

Was able to obtain the most complex model by removing BlockID as a random effect:

```{r}
ModelPostAnthByIns<- glmer(PostAnthBy8_28 ~ InsPool + Fert + standhgt + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelPostAnthByIns)
drop1(ModelPostAnthByIns, test = "Chi")
```

```{r}
ModelPostAnthByInsN<- glmer(PostAnthBy8_28 ~ InsPool + Fert + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelPostAnthByInsN)
drop1(ModelPostAnthByInsN, test = "Chi")
```

------------------------------------------------------------------------

## 1 Census Duration - InsExcl

There was no model that fit the duration under 1 Census Duration.

```{r}
ModelDurIns<- glm(X1CensFlrDuration ~ InsPool + Fert + standhgt , data = Phen1, family = binomial)
summary(ModelDurIns)
drop1(ModelDurIns, test = "Chi")
```

```{r}
ModelDurInsN<- glm(X1CensFlrDuration ~ InsPool + Fert, data = Phen1, family = binomial)
summary(ModelDurInsN)
drop1(ModelDurInsN, test = "Chi")
```

------------------------------------------------------------------------

## MaxHdsFlr - InsExcl

I removed all values that had NA as it means nothing if they do not ever have flowering heads.

```{r}
Phen2 <- subset(Phen1, MaxFlrBy8_26!='NA')
```

```{r}
ModelMaxFlrIns<- glm(MaxFlrBy8_26 ~ standhgt + Fert +InsPool, data = Phen2, family = binomial)
summary(ModelMaxFlrIns)
drop1(ModelMaxFlrIns, test = "Chi")
```

```{r}
ModelMaxFlrInsN<- glm(MaxFlrBy8_26 ~ Fert +InsPool, data = Phen2, family = binomial)
summary(ModelMaxFlrInsN)
drop1(ModelMaxFlrInsN, test = "Chi")
```

------------------------------------------------------------------------

# GLMM AgExcl

## Headby7/12 - AgExcl

```{r}
ModelHdByAg<- glmer(Headby7_12~ standhgt +Fert + AGPool +(1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelHdByAg)
drop1(ModelHdByAg, test = "Chi")
```

```{r}
ModelHdByAgN<- glmer(Headby7_12~ Fert + AGPool +(1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelHdByAgN)
drop1(ModelHdByAgN, test = "Chi")
```

## FlrBy8/26 - AgExcl

```{r}
ModelFlrByAg<- glmer(FlrBy8_26 ~ standhgt + AGPool + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelFlrByAg)
drop1(ModelFlrByAg, test = "Chi")
```

```{r}
ModelFlrByAg1<- glmer(FlrBy8_26 ~ Fert + AGPool + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelFlrByAg1)
drop1(ModelFlrByAg1, test = "Chi")
```

## PostAnthBy8/28 - AgExcl

```{r}
ModelPostAnthByAg<- glmer(PostAnthBy8_28 ~ AGPool + Fert + standhgt + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelPostAnthByAg)
drop1(ModelPostAnthByAg, test = "Chi")
```

```{r}
ModelPostAnthByAgN<- glmer(PostAnthBy8_28 ~ AGPool + Fert + (1|PlotID), data = Phen1, family = binomial,glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=300000)))
summary(ModelPostAnthByAgN)
drop1(ModelPostAnthByAgN, test = "Chi")
```

## 1CensDur - AgExcl

```{r}
ModelDurAg<- glm(X1CensFlrDuration ~ AGPool + Fert + standhgt, data = Phen1, family = binomial)
summary(ModelDurAg)
drop1(ModelDurAg, test = "Chi")
```

```{r}
ModelDurAgN<- glm(X1CensFlrDuration ~ AGPool + Fert, data = Phen1, family = binomial)
summary(ModelDurAgN)
drop1(ModelDurAgN, test = "Chi")
```

## MaxHdsFlr - AgExcl

```{r}
ModelMaxFlrAg<- glm(MaxFlrBy8_26 ~ standhgt + Fert + AGPool, data = Phen2, family = binomial)
summary(ModelMaxFlrAg)
drop1(ModelMaxFlrAg, test = "Chi")
```

```{r}
ModelMaxFlrAgN<- glm(MaxFlrBy8_26 ~ Fert + AGPool, data = Phen2, family = binomial)
summary(ModelMaxFlrAgN)
drop1(ModelMaxFlrAgN, test = "Chi")
```

------------------------------------------------------------------------

# Visualization

```{r}
library(ggplot2)
#Head by 7/12
InsHdBy <- as.data.frame(table(Phen1$InsPool,Phen1$Headby7_12))
InsHdBy <- setNames(InsHdBy,c("InsPool","Headby7_12","Freq"))
ggplot(InsHdBy, aes(x=InsPool, y=Freq, fill=Headby7_12))+
  geom_bar(stat="identity", position=position_dodge())+
  labs(title="Insect Exclusion on the Proportion of Any Head by July 12th")

#PostAnthBy 8/28
InsPostAnthBy <- as.data.frame(table(Phen1$InsPool,Phen1$PostAnthBy8_28))
InsPostAnthBy <- setNames(InsPostAnthBy,c("InsPool","PostAnthBy8_28","Freq"))
ggplot(InsPostAnthBy, aes(x=InsPool, y=Freq, fill=PostAnthBy8_28))+
  geom_bar(stat="identity", position=position_dodge())+
  labs(title="Insect Exclusion on the Proportion of Post Anthesis Heads by August 28th")

```

```{r}
ggplot(Phen1, aes(x=MaxPlantHgt, y=Headby7_12, color=Headby7_12))+
  geom_point()+
  stat_summary(fun="mean", geom="point", col="black")

ggplot(Phen1, aes(x=MaxPlantHgt, y=FlrBy8_26, color=FlrBy8_26))+
  geom_point()+
  stat_summary(fun="mean", geom="point", col="black")

ggplot(Phen1, aes(x=MaxPlantHgt, y=PostAnthBy8_28, color=PostAnthBy8_28))+
  geom_point()+
  stat_summary(fun="mean", geom="point", col="black")
```

```{r}
ModelPlantHgt<- lmer(MaxPlantHgt~Fert + InsPool +(1|PlotID), data = Phen1)
summary(ModelPlantHgt)
drop1(ModelPlantHgt, test = "Chi")



tapply(Phen1$MaxPlantHgt, Phen1$InsPool, mean)
tapply(Phen1$MaxPlantHgt, Phen1$InsPool, var)
qqnorm(resid(ModelPlantHgt))

tapply(Phen1$MaxPlantHgt, Phen1$Fert, mean)
tapply(Phen1$MaxPlantHgt, Phen1$Fert, var)
```


